From Net-Zero to Net-Negative: Assessing \hydrogen{} and \carbon{} Network Strategies in Europe

### 2023-11-25 report/submissions/journal/paper/paper.tex
rework abstract (still to be shortened significantly)

\begin{abstract}
    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles to enable the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, made from hydrogen and carbon dioxide: If carbonaceous fuel refineries are located at or near hydrogen production or import site, a \carbon{} network is needed to move \carbon{} from the cheapest carbon capture sites. If carbonaceous fuel refineries are located at or near carbon capture sites, a \hydrogen{} network is needed to transport \hydrogen{} from the cheapest hydrogen production sites. Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear.
    To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
    Our findings reveal that in a either-or scenario, the hydrogen network is superior to the carbon network as it enables the transportation cheap hydrogen inland to serve industrial processes and carbon capture and utilization at point sources. A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of low-cost hydrogen in energy-intensive regions such as industrial clusters.
    However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
    We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
    Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
    The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
\end{abstract}

@@

The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors. While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage. Hydrogen offers an efficient way to transport and store energy from different sources. Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested directly from the atmosphere using Direct Air Capture (DAC). Additionally, it can be stored in geological formations, a process known as carbon capture and sequestration (CCS). In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.

@@

However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems. A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CCU at the site of point-source capture. On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CCU.
In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,bec attiniCarbonDioxideCapture2022}. In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.. The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe... (extend on hydrogen literature)
The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway. To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
% However, the models are often limited with regard to both geographical scope and detail. While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.

@@

To supply carbon needed for Carbon Sequestration (CS) and Carbon Utilization (CU), the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities. The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility. This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Fig.~X). At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point. Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture. Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon. DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand. Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}. Biogas input is an expensive fuel which based on the endogenous modeling decisions is not used in large quantities. Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}. This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.

@@

System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process is cheapest. This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
Thus, total system costs in the Baseline model are highest, while the models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.

We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies. While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}). The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC. The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC. The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.

@@

The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure. The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers. This model shows a pronounced decrease in investments in solar generation and gas infrastructure. The \hydrogen{} network reduces DAC similarly to the \carbon{} network, but still produces 200 Mt of carbon from DAC. CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%. CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Fig.~\ref{fig:captureshare_line}). This variation highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks specific carbon capture potentials.

@@

The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost. The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with cheaply available green hydrogen. End-use primarily depends on the quality of renewable resources, and thus the availability of cheap hydrogen. In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered. Conversely, in Spain and the United Kingdom, which both have high renewable potentials and cheap hydrogen production costs, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.

In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce cheap hydrogen to make CU products (FT fuel, methanol, synthetic methane). In the absence of a \carbon{} network, sequestration sites first sequester all carbon from cheap point sources, and deploy additional DAC facilities at the locations with cheap renewables to capture the remaining necessary carbon for sequestration. Then, synfuel demand is met at the locations which offer the best mix of cheap carbon from point sources and cheap hydrogen from local renewables.
% Sequestered carbon is nearly always transported before

The \hydrogengrid{} on the other hand, transports cheap hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions. The main transportation routes go from Spain and the United Kingdom to Central Europe. In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources. This leads to a very decentralized synfuel production across the continent. Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula. In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above). In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network). Surplus hydrogen from lower FT fuel production is exported to Central Europe.

% Therefore, one can conclude that a \carbon{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources. A \hydrogen{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
These findings explain the advantage of the \hydrogen{} network in contrast to the \carbon{} network. Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon. While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with cheap renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production. The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with cheap hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with cheaper carbon available in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.

@@

Similar to the \hydrogengrid model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent. However, in the Hybrid model hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
The carbon network supplements this with smaller clusters of networks. In addition to a larger carbon grid cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes. These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea. These enable the system to transport cheaply available carbon from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}). This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks. Second-order effects can be seen in the shift in FT production compared to the \hydrogengrid model. Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland. A more detailed of differences between the \hydrogengrid and the Hybrid model can be found in Fig.~.

>>>>

\begin{abstract}
    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles.
    Hydrogen transport allows energy-intensive regions to import low-cost hydrogen. Carbon transport allows regions with carbon sinks such as carbon sequestration and synthetic fuel production to import low-cost carbon from distributed carbon capture sources.
    Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear. This is especially relevant for the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, which are made from both hydrogen and carbon dioxide, and in most locations require at least one of these feedstocks to be imported.
    To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
    Our findings reveal that when comparing either-or-models, a system with an hydrogen network provides lower system costs than a system with a carbon network. The former enables the transportation of hydrogen from centralized production sites with low-cost renewable energy in the Iberian Peninsula and the British Isles to serve distributed hydrogen demands across the European continent. The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for more expensive direct air capture. However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration sites. In all models, additional demand for synthetic fuels is met at sites with favorable renewable energy resources that enable both electrolysis and direct air capture at low costs.
    We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
    Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
    The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
\end{abstract}

@@

The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors. While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage. Hydrogen offers an efficient way to transport and store energy. Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested from the atmosphere using Direct Air Capture (DAC). Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS). In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Utilization (CU) strategies.

@@

However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems. A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CU at the site of point-source capture. On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CU.
In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}. In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.. The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe... (extend on hydrogen literature)
The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway. To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
% However, the models are often limited with regard to both geographical scope and detail. While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.

@@

To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities. The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility. This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Fig.~X). At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point. Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture. Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon. DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand. Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}. Biogas input is an expensive fuel which based on the endogenous modeling decisions is not used in large quantities. Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}. This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.

@@

System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process has the lowest cost. This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
Thus, total system costs in the Baseline model are highest, while the models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.

We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies. While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}). The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC. The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for expensive DAC. The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.

@@

The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure. The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers. This model shows a pronounced decrease in investments in solar generation and gas infrastructure. The \hydrogen{} network reduces DAC similarly to the \carbon{} network, but still produces 200 Mt of carbon from DAC. CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%. CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Fig.~\ref{fig:captureshare_line}). This variation highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks specific carbon capture potentials.

@@

The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost. The carbon grid transports low-cost carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with green hydrogen available at low costs. End-use primarily depends on the quality of renewable resources, and thus the availability of low-cost hydrogen. In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered. Conversely, in Spain and the United Kingdom, which both have high renewable potentials and low-cost hydrogen production, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.

In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane). In the absence of a \carbon{} network, sequestration sites first sequester all carbon from low-cost point sources, and deploy additional DAC facilities at the locations with low-cost renewables to capture the remaining necessary carbon for sequestration. Then, synfuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from local renewables.
% Sequestered carbon is nearly always transported before

The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions. The main transportation routes go from Spain and the United Kingdom to Central Europe. In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources. This leads to a very decentralized synfuel production across the continent. Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula. In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above). In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network). Surplus hydrogen from lower FT fuel production is exported to Central Europe.

% Therefore, one can conclude that a \carbon{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources. A \hydrogen{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
These findings explain the advantage of the \hydrogen{} network in contrast to the \carbon{} network. Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon. While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production. The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with low-cost hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with carbon available at lower costs in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.

@@

Similar to the \hydrogengrid model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent. However, in the Hybrid model hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
The carbon network supplements this with smaller clusters of networks. In addition to a larger carbon grid cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes. These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea. These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}). This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks. Second-order effects can be seen in the shift in FT production compared to the \hydrogengrid model. Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland. A more detailed of differences between the \hydrogengrid and the Hybrid model can be found in Fig.~.
